---
title: "Homework 5"
author: "Yifan Zhong"
date: "October 14, 2015"
output:  
      html_document:  
        keep_md: true  
---
#Load packages

```{r}
library(doBy)
```

Homework 5 Description

Load the Gross Domestic Product data for the 190 ranked countries in this data set: 
https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv 

Load the educational data from this data set: 
https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv 

Original data sources (if the links above don’t work): 
http://data.worldbank.org/data-catalog/GDP-ranking-table 
http://data.worldbank.org/data-catalog/ed-stats 

#download dataset
```{r}
gross_url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv "
education_url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv"
gross <- read.csv(gross_url,skip=4,na.strings=c("..",""),nrows=215)
gross <- gross[,c(1,2,4,5)]
colnames(gross) <- c("CountryCode","Ranking","Economy", "GDP")

education <- read.csv(education_url)
```

#Questions
#1	Match the data based on the country shortcode. How many of the IDs match?
189 data matched

#2	Sort the data frame in ascending order by GDP rank (so United States is last). What is the 13th country in the resulting data frame?
KNA : St. Kitts and Nevis   

```{r}
# Merge data
mergedata <- merge(gross, education, by="CountryCode",all=T)

#checking how many missing values
sapply(mergedata,function(var){sum(is.na(var))})

#delete observations missing GDP
data2 <- mergedata[!is.na(mergedata$GDP),]

#delete observations missing Income.Group
data2 <- data2[!is.na(data2$Income.Group),]

#checking missing values
sapply(data2,function(var){sum(is.na(var))})

#checking how many obs matched
dim(data2)

#sort data by GDP
data2$GDP <- as.numeric(gsub(",","",data2$GDP))
data2 <- data2[order(data2$GDP,na.last=F),]

# The 13th country in the resulting data frame
data2[13,c("CountryCode","Long.Name","Ranking", "GDP")]
```

#3	What are the average GDP rankings for the "High income: OECD" and "High income: nonOECD" groups? 
       

1 High income: nonOECD  -----  91.91304

2 High income: OECD ----- 32.96667

```{r}
q3 <- data2[which(data2$Income.Group %in% c("High income: OECD","High income: nonOECD" )),]
summaryBy(Ranking~Income.Group, data=q3)
```

#4	Cut the GDP ranking into 5 separate quantile groups. Make a table versus Income.Group. How many countries are Lower middle income but among the 38 nations with highest GDP?
5 countries are Lower middle income but among the 38 nations with highest GDP

```{r}
breakpoint <- quantile(data2$Ranking,probs=seq(0,1,length.out = 6))
data2$RankingGroup <- cut(data2$Ranking,breaks=breakpoint)
levels(data2$RankingGroup) <- c("first(top)", "second", "third", "fourth", "fifth")
table(data2$RankingGroup, data2$Income.Group)[,2:6]

#How many countries are Lower middle income but among the 38 nations with highest GDP?
sum(data2$Ranking<=38 & data2$Income.Group == "Lower middle income")
```

5	If you encounter NAs in your analysis, please delete those values and continue the analysis with the non-missing values. However, please include code to count the number of missing values for each variable used in the analysis.


Deliverable: Markdown file uploaded to GitHub containing the following 
•	Code for downloading, tidying, and merging data in a R Markdown file
•	Code to answer the four questions above (plus the answers) in the same R Markdown file
•	The file must be readable in GitHub – 5 points off if I have to download the file to read it!

In order to submit your assignment, send me the link to the file in GitHub via e-mail.

Rubric (25 points total):
Assignment is submitted via a link to GitHub that results in a readable file (0 or 1)
I can run the code on my computer with no modifications, except setting the working directory (0 – doesn’t run, 1 – partially runs, 2 – completely runs).
Correct answers for each of the first four questions (5 points each, 20 points total).
Correct estimate of NAs (2 points total).

